A Triple Origin for Twin Blue Stragglers in Close Binaries

### 2019-02-20 BSS7782.tex
edited tex file slightly for consistency with new figures

\begin{figure}[ht!]
\plotone{fig_doodle.eps}
\caption{Cartoon depiction of our proposed scenario for the formation
  of Binary 7782, specifically mass transfer from an evolved outer
  tertiary companion on to a compact inner binary via a circumbinary
  disk.  The outer tertiary component has mass m$_{\rm 3}$, whereas
  the inner binary components have masses m$_{\rm 1}$ and m$_{\rm 2}$.
  The inner and outer orbital separations are denoted by,
  respectively, a$_{\rm in}$ and a$_{\rm out}$.  The circularization
  radius of the accretion stream is denoted a$_{\rm c}$, as calculated
  via Equation~\ref{eqn:ac}, and marks the mean separation of the
  circumbinary disk.
\label{fig:fig1}}
\end{figure}

@@

\begin{figure}[ht!]
\plotone{fig_parameterspace.eps}
\caption{Parameter space in the P$_{\rm out}$-P$_{\rm in}$-plane
  allowed for the hypothetical outer tertiary orbit of Binary 7782
  before Roche-lobe overflow.  The solid diagonal black line shows the
  period corresponding to the circularization radius a$_{\rm c}$ for
  the mass transfer stream coming from the outer star (i.e., at the
  onset of mass transfer).  
%, assuming a mass ratio of q $=$ 0.7.
% and
%  component masses of m$_{\rm 1} =$ 0.9 M$_{\odot}$, m$_{\rm 2} =$ 1.1
%  M$_{\odot}$ and m$_{\rm 3} =$ 1.4 M$_{\odot}$.  
We assume initial component masses of m$_{\rm 1}$ = 1.1 M$_{\rm
  \odot}$ and m$_{\rm 2}$ = 0.9 M$_{\rm \odot}$ for the inner binary
components, and m$_{\rm 3}$ is computed for the outer tertiary
according to our assumed mass ratio (with our fiducial case
corresponding to q $=$ 0.7).  We assume completely conservative mass
transfer for this exercise, and a final mass for the outer tertiary of
0.6 M$_{\rm \odot}$ once it has become a WD.  The dashed diagonal
black line shows a rough criterion for dynamical stability in the
triple, approximately following \citet{1999ASIC..522..385M} (i.e.,
a$_{\rm in} <$ 0.1a$_{\rm out}$ is required for long-term dynamical
stability in equal-mass co-planar triples).  The vertical solid red
lines show the outer orbital periods corresponding to the hard-soft
boundary assuming central velocity dispersions of $\sigma =$ 1, 5 and
10 km s$^{-1}$.  The vertical dashed black lines show the maximum
outer orbital period P$_{\rm out}$ for which the outer tertiary
companion is Roche lobe-filling, assuming a stellar radius of R$_{\rm
  3} =$ 200 R$_{\odot}$ (which corresponds to the maximum stellar
radius reached on the AGB for the range of tertiary masses of interest
to us; see figure~\ref{fig:tertiarymass_vs_size}).  The horizontal
dashed red line shows the observed orbital period for Binary 7782,
using its observed orbital period and our assumed final inner
companion masses (i.e., m$_{\rm 1} =$ m$_{\rm 2} =$ 1.4 M$_{\odot}$).
Finally, the thick solid horizontal red line shows the parameter space
for P$_{\rm out}$ allowed after considering all of the aforementioned
criteria.
\label{fig:fig2}}
\end{figure}

@@

The stellar-evolution model, including the structure, temperature and
composition profiles are turned into a smoothed-particles
representation using the module {\tt StellarModelInSPH} in AMUSE (see
chapter 4 in \cite{AMUSE}).  We follow the same procedure as described
in \cite{2014MNRAS.438.1909D} for simulating the future of the triple
system $\chi$ Tau (HD 97131) in which the outer-most star overfills
its Roche lobe and transfers mass to an inner binary.  After
generating the hydrodynamical representation of the donor star we
replace the stellar core by a point mass to prevent the majority of
the resolution to be confined in the star's central regions.  In a
following step we relax the star using the hydrodynamics solver. This
relaxation process is realized in 100 steps during which we reduce the
velocity dispersion of individual SPH particles to a glasses structure
\citep[see, for example, \S\,3.3 on page 40
  in][]{1995LesHouchesSWhite}. During this procedure, the gaseous
envelope of the star tends to expand by about 20\%.  To determine the
radius of the evolving star we calculate Lagrangian radii and use the
distance to the stellar center which contains 90\% of its mass. From
this 90\% mass-radius relation we obtain the stellar radius and match
it with the Roche-lobe of the outer orbit.

@@

\begin{figure*}[ht!]
  \includegraphics[width=0.5\linewidth]{fig_BBSS_Lgas_M14Msun_A010au_t3360.pdf}
~  \includegraphics[width=0.5\linewidth]{fig_BBSS_gas_M14Msun_A010au_t3360.pdf}
\caption{Same as figure~\ref{fig:topview_at_t0} but at an age of $t
  \simeq 1091\,days$ after the start of the simulation. Left panel
  shows the equipotential surfaces of the triple overplotted with the
  gas distribution, the right panel shows just the gas and the stars
  as bullets.
\label{fig:topview_at_t1000day}}
\end{figure*}

>>>>

\begin{figure}[ht!]
%\plotone{fig_doodle.eps}
\includegraphics[width=\columnwidth]{fig_doodle.pdf}
\caption{Cartoon depiction of our proposed scenario for the formation
  of Binary 7782, specifically mass transfer from an evolved outer
  tertiary companion on to a compact inner binary via a circumbinary
  disk.  The outer tertiary component has mass m$_{\rm 3}$, whereas
  the inner binary components have masses m$_{\rm 1}$ and m$_{\rm 2}$.
  The inner and outer orbital separations are denoted by,
  respectively, a$_{\rm in}$ and a$_{\rm out}$.  The circularization
  radius of the accretion stream is denoted a$_{\rm c}$, as calculated
  via Equation~\ref{eqn:ac}, and marks the mean separation of the
  circumbinary disk.
\label{fig:fig1}}
\end{figure}

@@

\begin{figure}[ht!]
%\plotone{fig_parameterspace.eps}
\includegraphics[width=\columnwidth]{fig_parameterspace.pdf}
\caption{Parameter space in the P$_{\rm out}$-P$_{\rm in}$-plane
  allowed for the hypothetical outer tertiary orbit of Binary 7782
  before Roche-lobe overflow.  The solid diagonal black line shows the
  period corresponding to the circularization radius a$_{\rm c}$ for
  the mass transfer stream coming from the outer star (i.e., at the
  onset of mass transfer).  
%, assuming a mass ratio of q $=$ 0.7.
% and
%  component masses of m$_{\rm 1} =$ 0.9 M$_{\odot}$, m$_{\rm 2} =$ 1.1
%  M$_{\odot}$ and m$_{\rm 3} =$ 1.4 M$_{\odot}$.  
We assume initial component masses of m$_{\rm 1}$ = 1.1 M$_{\rm
  \odot}$ and m$_{\rm 2}$ = 0.9 M$_{\rm \odot}$ for the inner binary
components, and m$_{\rm 3}$ is computed for the outer tertiary
according to our assumed mass ratio (with our fiducial case
corresponding to q $=$ 0.7).  We assume completely conservative mass
transfer for this exercise, and a final mass for the outer tertiary of
0.6 M$_{\rm \odot}$ once it has become a WD.  The dashed diagonal
black line shows a rough criterion for dynamical stability in the
triple, approximately following \citet{1999ASIC..522..385M} (i.e.,
a$_{\rm in} <$ 0.1a$_{\rm out}$ is required for long-term dynamical
stability in equal-mass co-planar triples).  The vertical solid red
lines show the outer orbital periods corresponding to the hard-soft
boundary assuming central velocity dispersions of $\sigma =$ 1, 5 and
10 km s$^{-1}$.  The vertical dashed black line show the maximum
outer orbital period P$_{\rm out}$ for which the outer tertiary
companion is Roche lobe-filling, assuming a stellar radius of R$_{\rm
  3} =$ 200 R$_{\odot}$ (which corresponds to the maximum stellar
radius reached on the AGB for the range of tertiary masses of interest
to us; see figure~\ref{fig:tertiarymass_vs_size}).  The horizontal
dashed red line shows the observed orbital period for Binary 7782,
using its observed orbital period and our assumed final inner
companion masses (i.e., m$_{\rm 1} =$ m$_{\rm 2} =$ 1.4 M$_{\odot}$).
Finally, the thick solid horizontal red line shows the parameter space
for P$_{\rm out}$ allowed after considering all of the aforementioned
criteria.
\label{fig:fig2}}
\end{figure}

@@

The stellar-evolution model, including the structure, temperature and
composition profiles are turned into a smoothed-particles
representation using the module {\tt StellarModelInSPH} in AMUSE (see
chapter 4 in \cite{AMUSE}).  We follow the same procedure as described
in \cite{2014MNRAS.438.1909D} for simulating the future of the triple
system $\chi$ Tau (HD 97131) in which the outer-most star overfills
its Roche lobe and transfers mass to an inner binary.  After
generating the hydrodynamical representation of the donor star we
replace the stellar core by a point mass to prevent the majority of
the resolution to be confined in the star's central regions.  In a
following step we relax the star using the hydrodynamics solver. This
relaxation process is realized in 100 steps during which we reduce the
velocity dispersion of individual SPH particles to a glasses structure 
%\citep[see, for example, \S\,3.3 on page 40 in][]{1995LesHouchesSWhite}. 
(see, for example, \S\,3.3 on page 40 in White (1995)).  During this procedure, the gaseous
envelope of the star tends to expand by about 20\%.  To determine the
radius of the evolving star we calculate Lagrangian radii and use the
distance to the stellar center which contains 90\% of its mass. From
this 90\% mass-radius relation we obtain the stellar radius and match
it with the Roche-lobe of the outer orbit.

@@

\begin{figure*}[ht!]
  \includegraphics[width=0.5\linewidth]{fig_BBSS_Lgas_M14Msun_A010au_t3360.png}
~  \includegraphics[width=0.5\linewidth]{fig_BBSS_gas_M14Msun_A010au_t3360.png}
\caption{Same as figure~\ref{fig:topview_at_t0} but at an age of $t
  \simeq 1091\,days$ after the start of the simulation. Left panel
  shows the equipotential surfaces of the triple overplotted with the
  gas distribution, the right panel shows just the gas and the stars
  as bullets.
\label{fig:topview_at_t1000day}}
\end{figure*}
